### 2022-06-13 paper.tex
added imputation

Finally, we provide the above-mentioned methods to the scientific community in the form of a command line tool.
Our software, called \texttt{scbs}, allows the user to obtain a cell \times region matrix from raw methylation files using only few, simple shell commands
This means that users no longer need to write custom data-wrangling scripts, which ensures reproducibility across and makes the analysis of single-cell methylomes faster and easier.

@@

\subsection{PCA with imputation}
[...]

>>>>

Finally, we provide the above-mentioned methods to the scientific community in the form of a command line tool.
Our software, called \texttt{scbs}, allows the user to obtain a cell $\times$ region matrix from raw methylation files using only few, simple shell commands
This means that users no longer need to write custom data-wrangling scripts, which ensures reproducibility across and makes the analysis of single-cell methylomes faster and easier.

@@

Whenever a region is not covered by any read in a cell, the corresponding data entry in the input data matrix for the PCA will be missing. The standard approach to calculate PCA (commonly done using the IRLBA algorthm, \citep{Baglama2005}) is not suited to deal with missing data. We circumvent this issue by simply using the PCA itself to impute the missing value.

Let us write $A$ for the matrix to which the PCA is to be applied. In contrast to the main text, and in order to follow statistics rather than bioinformatcis convention, we consider here the features (here: regions) to be represented by the matrix rows rather than the columns. The matrix has already been centered, i.e., $\sum_i a_{ij}=0$. To establish notation, we remind the reader that performing a PCA on A means finding the singular value decomposition (SVD), $A=UDR^\top$, with $D$ diagonal and $U$ and $R$ orthonormal. The PCA scores are contained in $X=UD$, the loadings in $R$. To reconstruct the input data $A$ from the PCA representation, one may use $A=XR^\top$, i.e., $a_{ij}=\sum_r x_{ir} r_{jr}$, where the equation is exact if $r$ runs over all principal components and approximate if it is truncated to the leading ones.

Our iterative imputation strategy is now simply the following: we first replace all missing values in the row-centered input matrix $A$ with zeroes and perform the (truncated) PCA. Then, we use the PCA predictions for the missing values, i.e., the $a_{ij}=\sum_r x_{ir} r_{jr}$, as refined stand-ins for the missing values in $A$ and run the PCA once more. This can now be iterated utnil convergence, which is typically achieved after at most 4--5 rounds of PCA.
